Fixing Vulnerabilities Potentially Hinders Maintainability

### 2021-04-30 response/response_2/response.tex

\Us We thank the editor and reviewers for their valuable feedback and acceptance. In the 
following, we answer the questions raised by the reviewer 2.

\newreviewer{2}

\begin{revcomment}[2.1]
    What is meant by "20 person years"?  How would one obtain this measure?
\end{revcomment}

\Us A person-year is the equivalent to 12 person-months. A person-month is a standard measure of 
source code volume. It is the amount
of source code that one developer with average productivity could write in one
month. According to the "Building
Maintainable Software" book, 20 person-years of rebuild value 
is the equivalent to at most 175000 lines of code in Java. BCH uses person-years instead of volume 
because it facilitates the comparisons of source code volume between technologies
(different programming languages have different levels of of verbosity).

\begin{revcomment}[2.2]
    The bullet for automated tests.  Can you think of another work 
    instead of "contemplate"?  I'm not sure of what you are trying to say.
\end{revcomment}

\Us Thank you for pointing this out. We changed the work for ``include''.
This metric only considers unit testing, which is usually not applied for 
security testing but rather to functionality testing.

### 2021-04-30 paper/emse20.tex
small changes to response 2 and paper

2) The 
\emph{Automated Tests} guideline was also not considered since the 
tool does not consider two of the most important techniques to 
security testing: vulnerability scanning and penetration testing. 
Instead, it only considers unit testing.

@@

Our hypothesis is
that \emph{security patches hinder more software maintainability 
than regular changes}. 
We have seen, previously, a deterioration on software maintainability 
when patching vulnerabilities: $41.9\%$ ($406$) of patches suffered a 
negative impact, $38.7\%$ ($375$) of patches remained the same and $19.4\%$ 
($188$) of patches increased software maintainability. For regular changes,
when considering the size of the changes (\emph{size-basline}),
we observe that the maintainability decreases in $27.0\%$ ($262$) 
and increases in $30.5\%$ ($295$) of the cases. But in contrast to 
security patches, the maintainability of regular changes remains the 
same in $42.5\%$ ($412$) of the cases and perform regular
changes has a more positive impact than negative on  
maintainability. However, no statistical significance was 
obtained. Regular changes (\emph{random-baseline}), with no size 
restrictions, are less prone to hinder software maintainability than 
security changes. About $34.4\%$ ($333$) of the regular changes 
have negative impact on software maintainability---less than 
in the security patches. For the \emph{random-baseline}, 
statistical significance was retrieved ($p = 5.34$x$10^{-8}$).

@@

Overall, the results for both baselines, show that regular 
changes are less prone to hinder the software 
maintainability of open-source software.   
However, the \emph{size-baseline} integrates a larger 
number of cases with no impact on software maintainability. 
We manually inspected a total of $25$ cases from
that distribution of regular changes with no impact on maintainability, 
and, found that identifying regular changes with the same 
size as the security-related commit is limiting the type of regular 
commits being randomly chosen: input 
patches, variables or functions, type conversion (i.e., changes
with no impact on the software metrics analyzed by BCH).
We presume that this phenomenon lead
to the significant number of cases where there is no impact
on the software maintainability. On the other hand, 
identifying regular changes without any restrictions (\emph{random-baseline})
shows that regular changes have less negative impact 
on software maintainability compared with security patches
and that special attention should be given to security patches.

@@

\section{Study Implications}\label{sec:implications}
%
Our results show evidence that developers may have to reduce maintainability for the 
sake of security. We argue that developers should be able to patch and produce 
secure code without hindering the maintainability of their projects. But there are 
still concerns that need to be addressed and that this study brings awareness for:
%

\textit{\textbf{Follow Best Practices:}} Developers are not 
paying attention to some quality aspects of their solutions/patches, 
as seen in Figure~\ref{fig:guidelines}, ending up harming 
software maintainability. We argue that developers
should design and implement solutions that respect the 
limit bounds of branch points and function/module sizes that are 
recommended by best practices to avoid increasing the 
size, complexity and dependencies of their patches.
Developers should also keep function parameters below 
the recommended limit. It helps keep unit interfaces small 
and easy to use and understand.
Patterns such as \emph{Introduce Parameter Object} are useful 
to send information to a new function/class through an object
and keep the number of parameters small and the information well-organized.

Security patches also have a negative impact on the maintenance of 
software architecture. Maintaining the components independence and balance 
is important to make it easier to find the source code that 
developers want to patch/improve and to better understand
how the high-level componenet interact with others. Applying 
\emph{encapsulation} to hide implementation details and make the system 
more modular is a step forward not to hinder software architecture maintainability.

According to previous research, there is a correlation between 
code duplication and code smells~\cite{7476787}---duplicates are a source 
of regression bugs. BCH reports new code smells
for $15\%$ of the patches under study which supports previous research---$34\%$ of security patches 
introduce new problems~\cite{8819456,10.1145/3133956.3134072}. 
Developers should never reuse code by copying and pasting 
existing code fragments. Instead, they should create a method and call 
it every time needed. The \emph{Extract Method} refactoring 
technique solves many duplication problems. This makes spoting and 
solving the issue faster because you only need to fix one method instead
of multiple vulnerabilities. Clone detection tools such as CPD can help 
on locating duplicates.

@@

\textit{\textbf{Some Types of Vulnerabilities Need More Attention:}} 
Our study attempted at shedding light at the impact 
of different types of vulnerabilities on software maintainability.
Overall, all the CWEs under study present a negative impact
over $30\%$ on software maintainability. 
\emph{Cross-Site Scripting} (CWE-79) and \emph{Improper 
Restriction of Operations within the Bounds of a Memory Buffer} (CWE-119) 
are less prone to have impact on open-source software maintainability.
Developers should pay special attention to \emph{Improper Input Validation} (CWE-20), 
\emph{Information Exposure} (CWE-200), \emph{Missing Release of Memory after Effective Lifetime} 
(CWE-401) and \emph{Path Traversal} (CWE-22). However, 
more research should be performed to better understand 
the impact of each guideline on each CWE.

\textit{\textbf{Tools for Patch Risk Assessment Wanted:}}
	Design debt of one guideline can lead to severe impacts on the 
	software quality~\cite{10.1145/1985362.1985366}. Some software 
	producers consider security as a first class citizen while others 
	do not. As mentioned in previous work, security it is very important 
	and should be considered as a default feature~\cite{10.1145/2489828.2489830,kurilova2014wyvern,mcgraw2004software}. 
	However, the lack of experts and awareness to the ease of producing 
	vulnerable software leads companies to ship low quality software. 
	Providing automated tools to developers assess the risk of their 
  patches it is important to help companies shipping software of higher quality. 
  Bryan O’Sullivan, VP of Engineering at Facebook, advocated for new computer science
  risk models to detect vulnerabilities in scale and predict the level of security of the software under production in
  his talk “‘Challenges in Making Software Work at Scale”’ at FaceTAV’20.

	Tools like Better Code Hub can complement static analysis (e.g., 
  SonarQube, Codacy, ESLint, Infer, and more) to provide more information 
  to security engineers on the vulnerabilities per se and the risk of their patches
  ---alongside pull requests/code reviews.
  Static code analysis may be daunting due 
	to the number of rules and different effects
	on maintainability. BCH claims to use the top-$10$ of the guidelines
	with the highest effects on maintainability. Yet, static analysis 
	can be used to complement the BCH analysis by introducing
	the capability of vulnerability detection and support the prediction 
	and prevention of the risk of a patch of hindering maintainability.
%

\textit{\textbf{Computer Science Curricula Needs to be Updated:}}
	Computer Science curricula is not yet prepared to properly educate 
	students for maintainable security. Students should be exposed to 
	this problem to gain experience with it. Curricula should focus on 
	the production of secure and maintainable code and alert to the 
	trade-off between both. These matters should be discussed and 
	presented to students in software engineering courses.
	Universities should also encourage students to use code quality 
	analysis tools such as BCH or similar ones. These tools have a 
	great potential to make students aware of maintainability issues 
	and beginner mistakes (e.g., coding practices violation).
%

\textit{\textbf{Secure Programming Languages by Design:}} 
In general, our study 
shows that there is over $35\%$ of patches with negative impact 
on software maintainability per programming language. Developers implementing code in   
\texttt{C/C++}, \texttt{Ruby} and \texttt{PHP} should pay extra care 
to their solutions/patches. In addition, programming 
	languages should provide new design patterns to easily patch security weaknesses 
	without endangering maintainability. Ultimately, new programming languages, 
	both secure and maintainable by design, such as Wyvern~\cite{10.1145/2489828.2489830,kurilova2014wyvern}, 
	should be designed to help developers be less vulnerability prone when 
	writing and maintaining secure applications. One example is the need for 
	designing new authorization mechanisms since these are one of the most 
	complex security features to implement.

\section{Threats to Validity}\label{sec:threats}
%
This section presents the potential threats to validity of this study.
%

\textit{\textbf{Construct Validity:}} The formula to calculate the maintainability value ($M(v)$) was inferred based 
on the BCH's reports. The high amount of different projects and backgrounds 
may require other maintainability standards. However, BCH does use a 
representative benchmark of closed and open-source software projects to compute 
the thresholds for each maintainability guideline~\cite{Visser:2016:OREILLY,baggen2012}.
Maintainability is computed as the mean of all guidelines. Different software versions 
(vulnerable/fixed) of one vulnerability may have the same overall score
and still be affected by different guidelines. Therefore, we provide
an analysis per guideline and our results are all available on figshare for future reproductions and deeper analysis.

\textit{\textbf{Internal Validity:}} The security patches dataset provided by previous
work~\cite{Reis:2017:IJSSE} was collected based on the messages of GitHub
commits produced by project developers to classify the changes performed while 
patching vulnerabilities. This approach discards patches that were
not explicit in commits messages. We assume that patches were performed 
using a single-commit or several sequentially. The 
perspective that a 
developer may quickly perform a patch and later 
proceed to the refactor is not considered. We assume 
that all patches were only performed once. Depending 
on the impact of the vulnerability in the system, some 
vulnerabilities may have more urgency to be patched than 
others. For instance, a vulnerability performing a Denial-of-Service attack
which usually brings entire systems down may be more urgent to patch than a 
cross-site scripting vulnerability which usually does not have impact
on the systems execution but rather on the data accessibility.
We manually inspected $25.1\%$ of the security patches looking 
for floss-refactorings---$122$ from each dataset. We did 
find $23$ cases we argue to be floss-refactorings and toss them
in order to minimize the impact of this threat.

Baseline commits are retrieved randomly from 
the same project of the security patch.
This approach softens the differences
that may result from the characteristics of each project. However,
maintainability may still be affected by the developers' experience, coding
style, and software contribution policies which are not evaluated in this study.
Furthermore, this evaluation considers that $969$ regular commits---any kind
of commit---are enough to
alleviate random irregularities in the maintainability differences of the
baseline. 
%

\textit{\textbf{External Validity:}} The BCH tool uses private and open-source data to 
determine the thresholds for each guideline. We only analyze patches of open-source software.
Thus, our findings may not extend to private/non-open source software. Different programming 
languages may require different coding practices to address software safety. The 
dataset comprises more commits in Java, i.e., the dataset may not be representative 
of the population regarding programming languages. For both datasets, a manual validation 
of the message of the commits was performed. Only commits in English were considered. Thus, our approach does not consider 
patches in any other language but English.

@@

Many studies have investigated the relationship between patches and
software quality. Previous work focused on object-oriented metrics has evaluated the
impact of patches and exhibited proof that quantifying effectively the
impact of patches on maintainability may help to choose the appropriate
patch type~\cite{1167822}. In contrast to this work, Hegedus et
al.~\cite{HEGEDUS2018313} did not select particular metrics to assess the effect
of patches. Instead, statistical tests were used to find the metrics that
have the potential to change significantly after patches.

Researchers
performed a large-scale empirical study to understand the characteristics of security patches
and their differences against bug fixes~\cite{10.1145/3133956.3134072}.
The main findings were that security patches are smaller and less complex compared
to bug fixes and usually performed at function-level. Our study compares
the impact of security patches on software maintainability with the impact 
of regular changes.

Studying the evolution 
of maintainability issues during the development of Android apps, Malavolta et al. ($2018$)~\cite{8530041}
discovered that maintainability decreases over time. Palomba et al.
($2018$)~\cite{Palomba:2018:DIM:3231288.3231337} exhibits proof that code smells
should be carefully monitored by programmers since there is a high correlation
between maintainability aspects and proneness to changes/faults. In 2019, Cruz et 
al.~\cite{8919169} proposed a formula to calculate maintainability 
based on the BCH's guidelines and measured the impact of energy-oriented fixes 
on software maintainability. Recent work
proposed a new maintainability model to 
measure fine-grained code changes by adapting/extending the BCH model~\cite{8785997}.
Our work uses the same base model (SIG-MM), but considers a broader set of guidelines. 
Moreover, we solely focus on evaluating the impact of security patches on software maintainability.

@@

There are studies that investigated the impact of programming languages on software
quality~\cite{Ray:2014:LSS:2635868.2635922,Ray:2017:LSP:3144574.3126905}. The first
one shows that some programming languages are more buggy-prone than others. However,
the authors of the second one were not able to reproduce it and did not obtain any
evidence about the language design impact. 
Berger et al. ($2019$)~\cite{2019arXiv190110220B} tried to reproduce~\cite{Ray:2014:LSS:2635868.2635922,Ray:2017:LSP:3144574.3126905} 
and identified flaws that throw into distrust the 
previously demonstrated a correlation between programming language and software 
defects. Our work studies how security patches affect software quality based 
on the code maintainability analysis and provides shows that programming languages 
may have an impact on maintainability.

@@

This work presents an empirical study on the impact of $969$ security
patches on the maintainability of $260$ open-source projects. We leveraged
Better Code Hub reports to calculate maintainability based on a model proposed in 
previous work~\cite{Olivari:2018,8919169}. Results show evidence of a 
trade-off between security and maintainability, as $41.9\%$ of security patches 
yielded a negative impact. Hence,
developers may be hindering software maintainability while patching 
vulnerabilities. We also observe that some guidelines 
and programming languages are more likely to be affected than others. The implications 
of our study are that changes to codebases while patching 
vulnerabilities need to be performed with extra care; tools
for patch risk assessment should be integrate into the CI/CD pipeline; computer science
curricula needs to be updated; and, more secure programming languages are necessary.

As future work, the study can be extended in several directions: 
investigate which guidelines affect most the maintainability per
weakness; check if vulnerability patches are followed by patch
commits and how much time does it take to do it; 
expand our methodology with other software quality properties; 
validate these findings with closed/private
software; and, expand this analysis to other quality standards.

>>>>

2) The 
\emph{Automated Tests} guideline was also not considered since the 
tool does not include two of the most important techniques to 
security testing: vulnerability scanning and penetration testing. 
Instead, it only integrates unit testing.

@@

Our hypothesis is
that \emph{security patches hinder more software maintainability 
than regular changes}. 
We have seen, previously, a deterioration in software maintainability 
when patching vulnerabilities: $41.9\%$ ($406$) of patches suffered a 
negative impact, $38.7\%$ ($375$) of patches remained the same, and $19.4\%$ 
($188$) of patches increased software maintainability. For regular changes,
when considering the size of the changes (\emph{size-basline}),
we observe that the maintainability decreases in $27.0\%$ ($262$) 
and increases in $30.5\%$ ($295$) of the cases. But in contrast to 
security patches, the maintainability of regular changes remains the 
same in $42.5\%$ ($412$) of the cases, i.e., performing regular
changes has a more positive impact than negative on  
maintainability. However, no statistical significance was 
obtained. Regular changes (\emph{random-baseline}), with no size 
restrictions, are less prone to hinder software maintainability than 
security changes. About $34.4\%$ ($333$) of the regular changes hinder 
software maintainability---less than in the security patches. For the \emph{random-baseline}, 
statistical significance was retrieved ($p = 5.34$x$10^{-8}$).

@@

Overall, the results for both baselines show that regular 
changes are less prone to hinder the software 
maintainability of open-source software.   
However, the \emph{size-baseline} integrates a larger 
number of cases with no impact on software maintainability. 
We manually inspected a total of $25$ cases from
that distribution of regular changes with no impact on maintainability, 
and found that identifying regular changes with the same 
size as the security-related commit is limiting the type of regular 
commits being randomly chosen: input 
patches, variables or functions, type conversion (i.e., changes
with no impact on the software metrics analyzed by BCH).
We presume that this phenomenon lead
to the significant number of cases where there is no impact
on the software maintainability. On the other hand, 
identifying regular changes without any restrictions (\emph{random-baseline})
shows that regular changes have a less negative impact 
on software maintainability when compared to security patches
and that special attention should be given to security patches.

@@

\section{Study Implications}\label{sec:implications}
%
Our results show evidence that developers may have to reduce maintainability for the 
sake of security. We argue that developers should be able to patch and produce 
secure code without hindering the maintainability of their projects. But there are 
still concerns that need to be addressed and that this study brings awareness for:
%
\textit{\textbf{Follow Best Practices:}} Developers are not 
paying attention to some quality aspects of their solutions/patches, 
as seen in Figure~\ref{fig:guidelines}, ending up harming 
software maintainability. We argue that developers
should design and implement solutions that respect the 
limit bounds of branch points and function/module sizes that are 
recommended by best practices to avoid increasing the 
size, complexity, and dependencies of their patches.
Developers should also keep function parameters below 
the recommended limit. It helps keep unit interfaces small 
and easy to use and understand.
Patterns such as \emph{Introduce Parameter Object} are useful 
to send information to a new function/class through an object
and keep the number of parameters small and the information well-organized.

Security patches also harm the maintenance of software architecture. Maintaining the components independence and balance 
is important to make it easier to find the source code that 
developers want to patch/improve and to better understand
how the high-level components interact with others. Applying 
\emph{encapsulation} to hide implementation details and make the system 
more modular is a step forward not to hinder software architecture maintainability.

According to previous research, there is a correlation between 
code duplication and code smells~\cite{7476787}---duplicates are a source 
of regression bugs. BCH reports new code smells
for $15\%$ of the patches under study which supports previous research---$34\%$ of security patches 
introduce new problems~\cite{8819456,10.1145/3133956.3134072}. 
Developers should never reuse code by copying and pasting 
existing code fragments. Instead, they should create a method and call 
it every time needed. The \emph{Extract Method} refactoring 
technique solves many duplication problems. This makes spotting and 
solving the issue faster because you only need to fix one method instead
of multiple vulnerabilities. Clone detection tools such as CPD can help 
on locating duplicates.

@@

\textit{\textbf{Some Types of Vulnerabilities Need More Attention:}} 
Our study attempted to shed light on the impact of different types of vulnerabilities on software maintainability.
Overall, all the CWEs under study present a negative impact
over $30\%$ on software maintainability. 
\emph{Cross-Site Scripting} (CWE-79) and \emph{Improper 
Restriction of Operations within the Bounds of a Memory Buffer} (CWE-119) 
are less prone to have an impact on open-source software maintainability.
Developers should pay special attention to \emph{Improper Input Validation} (CWE-20), 
\emph{Information Exposure} (CWE-200), \emph{Missing Release of Memory after Effective Lifetime} 
(CWE-401) and \emph{Path Traversal} (CWE-22). However, 
more research should be performed to better understand the impact of each guideline on each CWE.

\textit{\textbf{Tools for Patch Risk Assessment Wanted:}}
  Design debt of one guideline can lead to severe impacts on the 
  software quality~\cite{10.1145/1985362.1985366}. Some software 
  producers consider security as a first-class citizen while others do not. As mentioned in previous work, security is critical and should be considered as a default feature~\cite{10.1145/2489828.2489830,kurilova2014wyvern,mcgraw2004software}. 
  However, the lack of experts and awareness of developers for security while producing/patching software leads companies to ship low-quality software. 
  Providing automated tools to developers to assess the risk of their patches is essential to help companies shipping software of higher quality. 
  Bryan O'Sullivan, VP of Engineering at Facebook, advocated for new computer science
  risk models to detect vulnerabilities in scale and predict the level of security of the software under production in his talk ```Challenges in Making Software Work at Scale''' at FaceTAV’20.

  Tools like Better Code Hub can complement static analysis (e.g., 
  SonarQube, Codacy, ESLint, Infer, and more) to provide more informatio

\section{Threats to Validity}\label{sec:threats}
%
This section presents the potential threats to the validity of this study.
%

\textit{\textbf{Construct Validity:}} The formula to calculate the maintainability value ($M(v)$) was inferred based 
on the BCH's reports. The high amount of different projects and backgrounds may require other maintainability standards. However, BCH does use a 
representative benchmark of closed and open-source software projects to compute the thresholds for each maintainability guideline~\cite{Visser:2016:OREILLY,baggen2012}.
Maintainability is computed as the mean of all guidelines. Different software versions 
(vulnerable/fixed) of one vulnerability may have the same overall score
and still be affected by different guidelines. Therefore, we provide
an analysis per guideline, and our results are all available on GitHub for future reproductions and deeper analysis.

\textit{\textbf{Internal Validity:}} The security patches dataset provided by previous
work~\cite{Reis:2017:IJSSE} was collected based on the messages of GitHub
commits produced by project developers to classify the changes performed while 
patching vulnerabilities. This approach discards patches that were
not explicit in commits messages. We assume that patches were performed using a single commit or several sequentially. The perspective that a 
developer may quickly perform a patch and later proceed to the refactor is not considered. We assume that all patches were only performed once. Depending on the impact of the vulnerability in the system, some vulnerabilities may have more urgency to be patched than others. For instance, a vulnerability performing a Denial-of-Service attack
that usually brings entire systems down may be more urgent to patch than a 
cross-site scripting vulnerability which generally does not have an impact
on the execution of the system but rather on the data accessibility.
We manually inspected $25.1\%$ of the security patches looking for floss-refactorings---$122$ from each dataset. We did 
find $23$ cases we argue to be floss-refactorings and toss them
to minimize the impact of this threat.

Baseline commits are retrieved randomly from 
the same project as the security patch.
This approach softens the differences
that may result from the characteristics of each project. However,
maintainability may still be affected by the developers' experience, coding
style, and software contribution policies which are not evaluated in this study.
Furthermore, this evaluation considers that $969$ regular commits---any kind
of commit---are enough to
alleviate random irregularities in the maintainability differences of the
baseline. 
%

\textit{\textbf{External Validity:}} The BCH tool uses private and open-source data to determine the thresholds for each guideline. We only analyze patches of open-source software.
Thus, our findings may not extend to private/non-open source software. Different programming languages may require different coding practices to address software safety. The dataset comprises more commits in Java, i.e., the dataset may not be representative of the population regarding programming languages. For both datasets, manual validation of the message of the commits was performed. Only commits in English were considered. Thus, our approach does not consider 
patches in any other language but English.

@@

Many studies have investigated the relationship between patches and
software quality. Previous work focused on object-oriented metrics has evaluated the
impact of patches and exhibited proof that quantifying the
impact of patches on maintainability may help to choose the appropriate
patch type~\cite{1167822}. In contrast to this work, Hegedus et
al.~\cite{HEGEDUS2018313} did not select particular metrics to assess the effect
of patches. Instead, statistical tests were used to find the metrics that
have the potential to change significantly after patches.

Researchers
performed a large-scale empirical study to understand the characteristics of security patches
and their differences against bug fixes~\cite{10.1145/3133956.3134072}.
The main findings were that security patches are smaller and less complex than bug fixes and are usually performed at the function level. Our study compares
the impact of security patches on software maintainability with the impact 
of regular changes.

Studying the evolution 
of maintainability issues during the development of Android apps, Malavolta et al. ($2018$)~\cite{8530041}
discovered that maintainability decreases over time. Palomba et al.
($2018$)~\cite{Palomba:2018:DIM:3231288.3231337} exhibits proof that code smells
should be carefully monitored by programmers since there is a high correlation
between maintainability aspects and proneness to changes/faults. In 2019, Cruz et 
al.~\cite{8919169} proposed a formula to calculate maintainability based on the BCH's guidelines and measured the impact of energy-oriented fixes 
on software maintainability. Recent work
proposed a new maintainability model to measure fine-grained code changes by adapting/extending the BCH model~\cite{8785997}.
Our work uses the same base model (SIG-MM) but considers a broader set of guidelines. 
Moreover, we solely focus on evaluating the impact of security patches on software maintainability.

@@

There are studies that investigated the impact of programming languages on software
quality~\cite{Ray:2014:LSS:2635868.2635922,Ray:2017:LSP:3144574.3126905}. The first
one shows that some programming languages are more buggy-prone than others. However,
the authors of the second one could not reproduce it and did not obtain any
evidence about the language design impact. 
Berger et al. ($2019$)~\cite{2019arXiv190110220B} tried to reproduce~\cite{Ray:2014:LSS:2635868.2635922,Ray:2017:LSP:3144574.3126905} 
and identified flaws that throw into distrust the 
previously demonstrated a correlation between programming language and software 
defects. Our work studies how security patches affect software quality based 
on the code maintainability analysis and provides shows that programming languages 
may have an impact on maintainability.

@@

This work presents an empirical study on the impact of $969$ security
patches on the maintainability of $260$ open-source projects. We leveraged
Better Code Hub reports to calculate maintainability based on a model proposed in previous work~\cite{Olivari:2018,8919169}. Results show evidence of a 
trade-off between security and maintainability, as $41.9\%$ of security patches yielded a negative impact. Hence,
developers may be hindering software maintainability while patching vulnerabilities. We also observe that some guidelines and programming languages are more likely to be affected than others. The implications of our study are that changes to codebases while patching vulnerabilities need to be performed with extra care; tools
for patch risk assessment should be integrated into the CI/CD pipeline; computer science
curricula need to be updated; and more secure programming languages are necessary.

As future work, the study can be extended in several directions: 
investigate which guidelines affect most the maintainability per
weakness; check if vulnerability patches are followed by new
commits and how much time does it take to do it; 
expand our methodology with other software quality properties; 
validate these findings with closed/private
software; and, expand this analysis to other quality standards.
